---
title: "HW_wk2"
output: html_document
date: "2022-10-14"
---
Here, I am importing the files using red_csv and here() which specifies where I want
```{r}
library(here)
library(sp)
library(sf)
library(tmap)
library(tmaptools)
library(RSQLite)
library(tidyverse)
library(rgdal)
library(readr)
library(dplyr)
library(OpenStreetMap)


ReportCardData1 <- read_csv("HW_wk2_data/Report_Card.csv",
                       locale = locale(encoding = "latin1"),
                       na = "NULL")
              

WashMap <- st_read(here::here("HW_wk2_data", "Washington_Counties_with_Natural_Shoreline___washsh_area", "Washington_Counties_with_Natural_Shoreline___washsh_area.shp"))

Datatypelist <- ReportCardData1 %>% 
  dplyr::summarise_all(class) %>%
  tidyr::pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
  
```
Above: I am looking to see that the columns I want are numeric or chr. Like to calculate the avg need a number

Below: Filtering rows to show all science students in all counties, then selecting columns, then checking the class of the data in each column
```{r}
library(janitor)
ReportCardData1 <- ReportCardData1 %>% 
  janitor::clean_names(., case="big_camel")


ReportCardData2 <- ReportCardData1 %>% 
  dplyr::filter(str_detect(TestSubject, "Science"),
         str_detect(GradeLevel, "All Grades"),
         str_detect(StudentGroup, "All Students"),
         str_detect(StudentGroupType, "All"),
         County != "Multiple") %>% distinct() %>% na.omit()

ReportCardData3 <- ReportCardData2 %>%
  janitor:: clean_names(., case="big_camel") %>%
  select(County, TestSubject, GradeLevel, StudentGroup, StudentGroupType, CountMetStandard, CountOfStudentsExpectedToTest)

rcd3_summary <- ReportCardData3 %>%
#I don't need to assign this to an obj, I can just say ReportCardData3 %>% but this is to see class and make sure I haven't accidentally changed the type of data in the columns
  summarise_all(class) %>% 
    pivot_longer(everything(),
               names_to="All_variables", 
               values_to="Variable_class")
```
Then I need to group into county, make a new column with the avg, but they are currently characters so I need to make them into numeric type

```{r}

ReportCardData3$passed <- as.numeric(ReportCardData3$CountMetStandard)
ReportCardData3$TotalStudents <- as.numeric(ReportCardData3$CountOfStudentsExpectedToTest)

grouped_studentsby_county <- ReportCardData3 %>%
  group_by(County) %>%
  summarise(passed_total = sum(passed), count_totalstudents = sum(TotalStudents)) %>% 
  mutate(county_avg = (passed_total/count_totalstudents)*100)%>%
  mutate(state_average = mean(county_avg))

arrange(grouped_studentsby_county)
```
Now I need to create a new tibble that says how far above/below the state avg it is, need to round

```{r}
compare_stateavg <- grouped_studentsby_county %>% 
  mutate(WAcompare = case_when(county_avg > 46.27107	 ~ "above state average", TRUE ~ "below state average"))
#Here, I am making a tibble saying if each county's avg, or county_avg, is below or above the state's average of 0.4741048

compare_stateavg1 <- compare_stateavg %>% 
  mutate(WAdiff = county_avg - 46.27107) %>% 
  group_by(WAcompare) %>%
  summarise(range=max(WAdiff)-min(WAdiff), count=n(), Average=mean(WAdiff))
#Here, I am making a tibble saying if each county's avg, or county_avg, is below or above the state's average of 0.4741048

head(compare_stateavg1)
#I just want to see what this new tibble looks like
```
```{r}
#This step is to make another tibble with all the counties and saying if they are above or below the state average, by how much, etc

compare_stateavg2 <- grouped_studentsby_county %>% 
  mutate(., WAdiff = county_avg-46.27107) %>%
  mutate(., across(where(is.numeric), round, 2))%>%
  mutate(., across(WAdiff, round, 1)) %>% 
  mutate(., WAcompare = case_when(county_avg >= 46.27107 ~ str_c("equal or above WA average by", WAdiff, "percent", sep=" "), TRUE ~ str_c("below below the state average by", WAdiff, "percent", sep=" "))) %>%
  mutate(WA_abovebelow = case_when(WAdiff < 0 ~ "below", WAdiff > 0 ~ "above", WAdiff == 0 ~ "equal"))

compare_stateavg3 <- compare_stateavg2%>%
  #select(County) %>% 
  group_by(WA_abovebelow)%>%
  summarise(count=n())

arrange(compare_stateavg2)

#HOW DO I INCLUDE SUMMARISE HERE?
```

MAPPING, yay :)

```{r}
#setup
#install.packages("maptools")
#install.packages(c("classInt", "tmap"))
#install.packages(c("RColorBrewer", "sp", "rgeos", 
#                   "tmaptools", "sf", "downloader", "rgdal", 
#                   "geojsonio"))

library(maptools)
library(RColorBrewer)
library(classInt)
library(sp)
library(rgeos)
library(tmap)
library(tmaptools)
library(sf)
library(rgdal)
library(geojsonio)

#now read in geo-data!! Shape file was read in at the beginning of this hw

#qtm(WashMap)

#making more detailed map, WHAT TYPE OF JOIN DO I NEED
clean_names(WashMap)

ReportCardData_Map <- WashMap %>% 
  clean_names() %>% 
  left_join(., compare_stateavg2, by = c("countylabe" = "County"))

library(tmap)
qtm(ReportCardData_Map, fill = "WA_abovebelow", fill.palette="-Blues")

library(tmap)
qtm(ReportCardData_Map, fill = "WAdiff", fill.palette="-Greens")

#I need to figure out the column name stuff

#Add base map - they did two maps, one on open street map, but I don't need to do that and idk if I want to do it on open street map anyway... so I need to join the WAdiff column to the map by county as the connector then just plot it below, but I need to define tmapWA so don't delete that part. Ask Phil tho if I even need that part because maybe I can just use ReportCardData_Map, like does th tmap part of tmapWA do anything?

```



```{r}
#install.packages("shinyjs")
#palette_explorer()
tmapWA <- ReportCardData_Map %>%
  st_bbox(.) %>%
  tmaptools::read_osm(., type = "osm", zoom = NULL)

tmap_mode("plot")

tm_shape(tmapWA)+
tm_rgb()+
tm_shape(ReportCardData_Map) + 
tm_polygons("WA_abovebelow", 
        style="jenks",
        palette="YlOrBr",
        midpoint=NA,
        title="Percent above/below state avg",
        alpha = 0.5) + 
  tm_compass(position = c("left", "bottom"),type = "arrow") + 
  tm_scale_bar(position = c("left", "bottom")) +
  tm_layout(title = "Percent above/below state avg", legend.position = c("right", "bottom"))
```

